---
title: "Homework2"
author: "Jon Campbell"
format: pdf
editor: visual
---

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(knitr)
library(PSweight)
```

## Part 1

### Question 1

```{r}
hsr <- read.table("data/hw2/HSR.txt", header = TRUE)
#p_z = P(Y(z)==1)
#target estimand is ATE p_1 - p_2
X <- apply(hsr[,c(1,3,4,5,6,7,8,12)], MARGIN = 2, factor)
Z <- factor(hsr$pg)
Y <- hsr$i_aqoc
```

#### a)

```{r}
ps_formula <- as.factor(pg) ~ i_age + as.factor(i_sex) + as.factor(i_race) + as.factor(i_educ) + as.factor(i_insu) + as.factor(i_drug) + as.factor(i_seve) + com_t + pcs_sd + mcs_sd

prop_scores <- glm(ps_formula, family = "binomial"
                   , data = hsr)$fitted.values

prop_score_summary <- SumStat(ps.formula = ps_formula
                            , weight = c('IPW','overlap')
                            ,data = hsr)
```

```{r}
#Estimated propensity Scores
cat(prop_scores)
```

```{r}
#checking balance via ASD
plot(prop_score_summary, metric = "ASD")
```

From the love plot we see that most of the covariates have an ASD above 0.1 when unweighted but perform much better with IPW. mcs_sd and severity2 are slightly above 0.1 and race2 is around 0.35 under IPW. The covariates are perfected balanced with overlap weighing.

```{r}
#Inverse Probability Weights 
IPW_weights <- ifelse(Z == 2, 1/prop_scores,1/(1-prop_scores))

#Normalized IPW 
IPW_weights_norm <- numeric(length(IPW_weights))
IPW_weights_norm[Z==2] = IPW_weights[Z==2]/sum(IPW_weights[Z==2])
IPW_weights_norm[Z==1] = IPW_weights[Z==1]/sum(IPW_weights[Z==1])
ggplot(data.frame(weights = IPW_weights
                  ,high=as.factor(ifelse(IPW_weights>15,1,0)))
       , aes(x = weights, fill = high)) +
  geom_histogram(bins = 20) +
  labs(title = "Distribution of IPW", x = "IPW")
```

Looking at he distribution of IPW there are `r sum(IPW_weights>17)` observations with high weights relative to the sample.

#### b)

```{r}
tibble(
  prop_score = prop_scores,
  trt = Z
) |>
  ggplot(aes(x = prop_score, fill = trt)) +
  geom_density(alpha = 0.5)
plot(prop_score_summary, metric = "ASD")
```

From the density plot of propensity scores by group, there is good overlap. The love plot shows good balance with overlap weighting and IPW weighting.

#### c)

```{r}
unweighted <- prop_score_summary$unweighted.sumstat[,1:2]
colnames(unweighted) <- c("unweighted.trt1", "unweighted.trt2")

ipw_weighted <- prop_score_summary$IPW.sumstat[,1:2]
colnames(ipw_weighted) <- c("IPW.trt1", "IPW.trt2")

overlap_weighted <- prop_score_summary$overlap.sumstat[,1:2]
colnames(overlap_weighted) <- c("overlap.trt1", "overlap.trt2")

tab1 <- cbind(unweighted, ipw_weighted, overlap_weighted) |> round(4)
tab1 |>
  kable()
```

#### d)

```{r}
Z <- ifelse(Z == 1,0,1)
ps.hsr <- as.factor(pg) ~ i_age + as.factor(i_sex) + as.factor(i_race) + as.factor(i_educ) + as.factor(i_insu) + as.factor(i_drug) + as.factor(i_seve) + com_t + pcs_sd + mcs_sd

ate.ipw.norm <- PSweight(ps.formula = ps.hsr, yname = "i_aqoc", data = hsr
                    , weight = "IPW")
ato.overlap <- PSweight(ps.formula = ps.hsr, yname = "i_aqoc", data = hsr
                    , weight = "overlap")
ate.ipw.nonnorm <- summary(lm(i_aqoc ~ pg, data = hsr, weights = IPW_weights))$coefficients[2,1:2]
results.ate <- rbind(summary(ate.ipw.norm, type = "DIF", contrast = c(-1,1))$estimates[,1:2],
      summary(ato.overlap, type = "DIF", contrast = c(-1,1))$estimates[,1:2],
      ate.ipw.nonnorm)
rownames(results.ate) <- c("ATE IPW Normalized", "ATO Overlap", "ATE NonNormalized")
results.ate |> round(4)

```



### Question 2

#### a)

```{r}
ps_formula <- as.factor(pg) ~ i_age + as.factor(i_sex) + as.factor(i_race) + as.factor(i_educ) + as.factor(i_insu) + as.factor(i_drug) + as.factor(i_seve) + com_t + pcs_sd + mcs_sd

prop_scores <- glm(ps_formula, family = "binomial"
                   , data = hsr)$fitted.values

hsr_subsetted <- subset(hsr, i_sex == 1)
hsr_subsetted$ipw_weights <- ifelse(hsr_subsetted$pg == 1, 1/prop_scores,1/(1-prop_scores))
att.subgroup <- summary(lm(i_aqoc ~ pg + i_age + i_sex + as.factor(i_race) + as.factor(i_educ) + as.factor(i_insu) + as.factor(i_drug) + as.factor(i_seve) + com_t + pcs_sd + mcs_sd, weights = ipw_weights, data = hsr_subsetted))$coefficients[2,1:2]
att.subgroup
```

#### b)

```{r}
ps_formula <- as.factor(pg) ~ i_age + as.factor(i_race) + as.factor(i_educ) + as.factor(i_insu) + as.factor(i_drug) + as.factor(i_seve) + com_t + pcs_sd + mcs_sd

prop_scores <- glm(ps_formula, family = "binomial"
                   , data = hsr)$fitted.values

hsr_subsetted <- subset(hsr, i_sex == 1)
hsr_subsetted$ipw_weights <- ifelse(hsr_subsetted$pg == 1, 1/prop_scores,1/(1-prop_scores))

att.subgroup <- summary(lm(i_aqoc ~ pg + i_age + i_sex + as.factor(i_race) + as.factor(i_educ) + as.factor(i_insu) + as.factor(i_drug) + as.factor(i_seve) + com_t + pcs_sd + mcs_sd, weights = ipw_weights, data = hsr_subsetted))$coefficients[2,1:2]
att.subgroup
```


#### c)

```{r}
hsr_subsetted <- subset(hsr, i_sex == 1)
ps.hsr <- as.factor(pg) ~ i_age + as.factor(i_race) + as.factor(i_educ) + as.factor(i_insu) + as.factor(i_drug) + as.factor(i_seve) + com_t + pcs_sd + mcs_sd


summary(PSweight(ps.formula = ps.hsr, yname = "i_aqoc", data = hsr_subsetted
                    , weight = "IPW"))$estimates[,1:2]
```



### Question 3

#### a)

```{r}

```

#### b)

```{r}

```

#### c)

```{r}

```

#### d)

```{r}

```

#### e)

```{r}

```

## Part 2

```{r}
brscn <- read.table("data/hw2/brscn.txt", header = TRUE)
```

### Question 1

```{r}

```


